The Pair Distribution Function of the Inhomogeneous Hard-Sphere Fluid

### 2014-01-18 papers/pair-correlation/paper.tex

%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  We have developed an approximation for the pair distribution
  function of hard spheres, which is based on the average value of the
  pair distribution function at contact described in
  Section~\ref{sec:contact}.  Our pair distribution is constructed so
  as to be computable with only single-site convolutions and to
  accurately reproduce the averaged distribution function at contact.
  This results in a functional that is efficient to compute, and
  predicts the pair distribution function as well as the most accurate
  existing analytical approximation.
\end{abstract}

@@

The standard approach in liquid state theory is to treat a liquid as a
hard-sphere reference fluid with attractive interactions that are
treated perturbatively~\cite{hansen2006theory}.  Recently there have
been considerable advances in the development of classical density
functionals have been based on the
application of these theoretical approaches to inhomogeneous
densities~\cite{ jain2007modified, gloor2007prediction, gross2009density,
  %felipe2001examination, gloor2002saft,
  %gloor2004accurate, clark2006developing,
  kahl2008modified,
  % yu2002fmt-dft-inhomogeneous-associating,
  %fu2005vapor-liquid-dft, bryk2006density,
  hughes2013classical,
  %segura1998comparison, felipe2001examination,
  %gloor2002saft, gloor2004accurate, fu2005vapor-liquid-dft,
  bryk2006density, clark2006developing,
  kahl2008modified, gross2009density,
  %yu2002fmt-dft-inhomogeneous-associating,
  sundararaman2012computationally, marshall2013density}.  The
perturbation theory treatment of intermolecular interactions relies on
the pair distribution function of the reference fluid:
$g^{(2)}(\rr_1,\rr_2)$.  Unlike the radial distribution function of a
homogeneous fluid, there does not currently exist a tractable form for
the pair distribution function of an inhomogeneous hard-sphere fluid,
suitable for use in constructing a density
functional~\cite{gloor2007prediction, jain2007modified}.

@@

The first canonical inhomogeneous configuration for the hard-sphere
fluid is the system consisting of a hard spheres at a hard wall.  In
1986, Plischke and Henderson solved the pair distribution function of
this system using integral equation theory under the Percus-Yevick
approximation~\cite{plischke1986pair}.  Ten years later,
G{\"o}tzelmann \emph{et al.} found the pair distribution function near
a hard wall using classical DFT by finding the direct correlation
function by taking a functional derivative of the free energy
functional and then solving the Ornstein-Zernicke
equation~\cite{gotzelmann1996structure}.  The approach of solving the
Ornstein-Zernicke equation for an inhomogeneous system is
computationally challenging~\cite{paul2003variational}.

@@

Approaches that do not diverge use an average of the density over some
volume. Fischer and Methfessel~\cite{fischer1980born} introduce the
approximation
\begin{align}
  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, n_3\left(\tfrac12
  (\rr_1+\rr_2)\right)\right)
  \label{eq:fischer}
\end{align}
where $n_3$ is one of the fundamental measures defined in Fundamental
Measure Theory~\cite{rosenfeld1989free}, which is an integral of the
density over a spherical volume:
\begin{align}
  n_3(\rr) = \int n(\rr')\Theta(\tfrac12 \sigma - |\rr-\rr'|) d\rr'
\end{align}
where $\sigma$ is the hard-sphere diameter.  Equation~\ref{eq:fischer}
is computationally awkward, because it relies on the midpoint
$\tfrac12(\rr_1+\rr_2)$.  Moreover, the approach of Fischer and
Methfessel is intended to the pair distribution function only at contact,
when the distance between $\rr_1$ and $\rr_2$ is a hard-sphere
diameter.

For these reasons Sokolowski and Fischer modified this approach to use
averages centered on the two points $\rr_1$ and $\rr_2$:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12},
  \tfrac12(\bar{n}(\rr_1)+\bar{n}(\rr_2))\right)
  \label{eq:sokolowski}
\end{align}
where
\begin{align}
  \bar{n}(\rr) = \frac{3}{4\pi (0.8\sigma)^3}\int n(\rr')\Theta(0.8\sigma - |\rr-\rr'|) d\rr'
\end{align}
is the density averaged over a sphere with diameter
$0.8\sigma$~\cite{sokolowski1992role}.  The value 0.8 in this formula
was arrived at by fitting to Monte Carlo simulation.  Although
Eq.~\ref{eq:sokolowski} has the advantages of only involving density
averages at the points at which the pair distribution function is
desired, cannot be written as a single-site convolution, since the
convolution kernel depends on both points.

@@

We construct our function's form in terms of an expansion of powers of
$h_{\sigma} = g_{\sigma}-1$, ensuring that it approaches $g_{\sigma}$
in the contact distance limit, leaving our function with tunable
constant perameters.  We run a simple monte-carlo simulations to
calculate values of the homogeneous radial distribution function at
varying distance perameter $r$.  We then tune our function's constant
perameters by matching our function's results against the monte-carlo
data, using a least squares fit method.  During this process it is
simplest to use an analytic expression for $g_{\sigma}$ in our
function and so we use the well known and very successful Carnahan
Starling equation to calculate $g_{\sigma}$.  This is appropriate
since that our published $g_{\sigma}$ approaches the Carnahan Starling
version in the homogeneous limit and it's this limit, along with the
contact limit, that we would like our constructed radial distribution
function to get right.  After this process we have an appropriate
radial distribution function that is dependent only on $r$ and
$g_{\sigma}$.

The function has the form:
\begin{multline}
  g(|\rr|) = 1 + g_{\sigma}e^{-x_0(r-\sigma)} \\+ x_1g_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
  -x_4g_{\sigma}^2\sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
\end{multline}
The finalalized form of our pair correlation function is:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
\end{align}

@@

In contrast to the previous work, which are based on the homogeneous
radial distribution function $g(r,\eta)$ expressed as a function of
distance and packing fraction, we have constructed our separable
approximation for the radial distribution function as $g(r,
g_\sigma)$, which is a function of the radius $r$ and its value at
contact $g_\sigma$.  One can convert between these two representation
using the Carnahan-Starling formula for $g_\sigma(\eta)$
\begin{equation}
  g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}.
\end{equation}
The form $g(r, g_\sigma)$ is more convenient for treating the
inhomogeneous case, because we can directly compute the averaged value
of $g_\sigma$ for an inhomogeneous system~\cite{schulte2012using}.

We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
\begin{equation}
  g(\sigma, g_\sigma) = g_\sigma.
\end{equation}
A second constraint applies to the integral of the total correlation
function
\begin{align}
  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
\end{align}
where $h(r) = g(r) - 1$ is the total correlation function.
A final constraint is that the slope of the radial distribution
function at contact is given to a very good approximation by
\begin{align}
  g'(\sigma) \approx - h_\sigma g_\sigma
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function.
Under these three constraints, we applied a least-squares fit to
simulation data for the radial distribution function at packing
fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
distribution function is given by
\newcommand\kappaa{\kappa_0}
\newcommand\kappab{\kappa_1}
\newcommand\kappac{\kappa_2}
\begin{multline}
  h(r,g_\sigma) = 1 + h_\sigma e^{-\kappaa z} \\
  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
  + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
      \eta^4}\frac{1}{n}
    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
\end{multline}
\begin{equation}
  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
  - \frac{4\pi}{3} \sigma^3
\end{equation}
\begin{equation}
  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
\end{equation}
\begin{equation}
  I_2 = \frac{8\pi (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
\end{equation}
\begin{equation}
  z = r - \sigma
\end{equation}
\davidsays{Can someone check the above equations for correctness
  against the python code, in plot-ghs2.py.  We'll want to
  ensure the C++ code agrees with this as well.}
This function includes three fitted parameters, $\kappaa$, $\kappab$,
and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
figure, it is apparent how the on the integral of $h(r)$ from
Eq.~\ref{eq:total-constraint} results in cancelling errors at large
distances.  The maximum error in $g(r)$ at packing fractions of less
than 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr$.

@@

Figure~\ref{fig:pair-distribution-1}b
and~\ref{fig:pair-distribution-3}b show the pair distribution function
as plotted along a path illustrated in
Fig.~\ref{fig:pair-distribution-1}a.  These plots compare this work
with Monte Carlo results, as well as the approximations of Sokolowski
and Fischer~\cite{sokolowski1992role} and that of Fischer and
Methfessel~\cite{fischer1980born} at packing fractions of 0.1 and 0.3.
The approach of Fischer and Methfessel is only valid when the two
spheres are in contact, and is therefore only plotted on that segment
of the path.  As an input to both previous approximations we use the
hard sphere radial distribution function found using Monte Carlo
simulation, interpolated as necessary.  We find that both previous
approximations to the pair distribution function predict the
qualtitative angular dependence of the pair distribution function at
contact better than this work.  However, in each case the pair
distribution function has a systematic error at contact---either too
high or too low.  Thus while our approximation is smoother than either
of the existing approaches, its errors will have a tendancy to cancel
when used in a perturbation expansion.  At higher densities, the
approximation of Fischer and Methfessel requires evaluating the radial
distribution function at densities significantly higher than the
freezing density, which poses numerical difficulties when using the
radial distribution function from simulation.  When the two points are
more than a radius away from contact, we find that any of these
approaches gives a reasonable prediction, with our approach
underestimating the oscillations in $g^(2)$, as expected based on our
fit in Fig.~\ref{fig:radial-distribution}.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
  \caption{The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
    shows the triplet distribution function, with white corresponding
    to the contact value of the \emph{pair} distribution function.
    The plot on the right shows the value of $g^{(3)}$ along the path
    shown as a dotted line on the plot to the
    left.}\label{fig:triplet-contact-distribution-3}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
  \caption{The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
    shows the triplet distribution function, with white corresponding
    to the contact value of the \emph{pair} distribution function.
    The plot on the right shows the value of $g^{(3)}$ along the path
    shown as a dotted line on the plot to the
    left.}\label{fig:triplet-contact-distribution-4}
\end{figure}

@@

Figures~\ref{fig:triplet-contact-distribution-3}-\ref{fig:triplet-inbetween-distribution-4}
compare the triplet distribution function computed using our
approximation to $g^{(2)}$ with results from Monte Carlo simulation.
In each figure $\rr_1$ and $\rr_2$ are held fixed, while the third
position $\rr_3$ is varied.  The test-particle position in each case
is $\rr_1$, which is on the left-hand side of the figure.  In the
color plot on the left, the first and second spheres are displayed in
grey.  On the bottom half of each left-hand plot is the triplet
correlation function as computed using the approximation presented in
this work.  The top half displays the triplet correlation function
from Monte Carlo.  On the right, we display

@@

Figure~\ref{fig:dadz-delta} shows the results for the sticky
hard-sphere fluid.  Our pair distribution function is constructed to
produce this result exactly, provided the averaged pair distribution
function at contact from Ref.~\citenum{schulte2012using} is exact.  As
expected, we see excellent agreement with the Monte Carlo simulation
results, while the approximations of Fischer and Sokolowski each show
deviations near the interface.  Figure~\ref{fig:dadz-square-well}
shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
This potential stresses our approximation to the radial distribution
function, since the edge of the attraction occurs at the dip in the
radial distribution function at this density, which poor results for
the bulk fluid.  We see that over these length scales, Sokolowski's
approximation is very accurate.  Finally, in the case of the power-law
attraction both approximations work very well.

>>>>

%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  We have developed an approximation for the pair distribution
  function of hard spheres, which is based on the average value of the
  pair distribution function at contact described in
  Section~\ref{sec:contact}.  Our pair distribution function is constructed so
  as to be computable with only single-site convolutions and to
  accurately reproduce the averaged distribution function at contact.
  This results in a functional that is efficient to compute, and
  predicts the pair distribution function as well as the most accurate
  existing analytical approximation.
\end{abstract}

@@

The standard approach in liquid state theory is to treat a liquid as a
hard-sphere reference fluid with attractive interactions that are
treated perturbatively~\cite{hansen2006theory}.  Recently there have
been considerable advances in the development of classical density
functionals based on the
application of these theoretical approaches to inhomogeneous
densities~\cite{ jain2007modified, gloor2007prediction, gross2009density,
  %felipe2001examination, gloor2002saft,
  %gloor2004accurate, clark2006developing,
  kahl2008modified,
  % yu2002fmt-dft-inhomogeneous-associating,
  %fu2005vapor-liquid-dft, bryk2006density,
  hughes2013classical,
  %segura1998comparison, felipe2001examination,
  %gloor2002saft, gloor2004accurate, fu2005vapor-liquid-dft,
  bryk2006density, clark2006developing,
  kahl2008modified, gross2009density,
  %yu2002fmt-dft-inhomogeneous-associating,
  sundararaman2012computationally, marshall2013density}.  The
perturbation theory treatment of intermolecular interactions relies on
the pair distribution function of the reference fluid:
$g^{(2)}(\rr_1,\rr_2)$.  Unlike the radial distribution function of a
homogeneous fluid, there does not currently exist a tractable form for
the pair distribution function of an inhomogeneous hard-sphere fluid,
suitable for use in constructing a density
functional~\cite{gloor2007prediction, jain2007modified}.

@@

The first canonical inhomogeneous configuration for the hard-sphere
fluid is the system consisting of a hard sphere at a hard wall.  In
1986, Plischke and Henderson solved the pair distribution function of
this system using integral equation theory under the Percus-Yevick
approximation~\cite{plischke1986pair}.  Ten years later,
G{\"o}tzelmann \emph{et al.} found the pair distribution function near
a hard wall using classical DFT by finding the direct correlation
function by taking a functional derivative of the free energy
functional and then solving the Ornstein-Zernicke
equation~\cite{gotzelmann1996structure}.  The approach of solving the
Ornstein-Zernicke equation for an inhomogeneous system is
computationally challenging~\cite{paul2003variational}.

@@

Approaches that do not diverge use an average of the density over some
volume. Fischer and Methfessel~\cite{fischer1980born} introduce the
approximation
\begin{align}
  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, n_3\left(\tfrac12
  (\rr_1+\rr_2)\right)\right)
  \label{eq:fischer}
\end{align}
where $n_3$ is one of the fundamental measures defined in Fundamental
Measure Theory~\cite{rosenfeld1989free}, which is an integral of the
density over a spherical volume:
\begin{align}
  n_3(\rr) = \int n(\rr')\Theta(\tfrac12 \sigma - |\rr-\rr'|) d\rr'
\end{align}
where $\sigma$ is the hard-sphere diameter.  Equation~\ref{eq:fischer}
is computationally awkward, because it relies on the midpoint
$\tfrac12(\rr_1+\rr_2)$.  Moreover, the approach of Fischer and
Methfessel is intended to approximate the pair distribution function only
at contact, when the distance between $\rr_1$ and $\rr_2$ is a hard-sphere
diameter.

For these reasons Sokolowski and Fischer modified this approach to use
averages centered on the two points $\rr_1$ and $\rr_2$:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12},
  \tfrac12(\bar{n}(\rr_1)+\bar{n}(\rr_2))\right)
  \label{eq:sokolowski}
\end{align}
where
\begin{align}
  \bar{n}(\rr) = \frac{3}{4\pi (0.8\sigma)^3}\int n(\rr')\Theta(0.8\sigma - |\rr-\rr'|) d\rr'
\end{align}
is the density averaged over a sphere with diameter
$0.8\sigma$~\cite{sokolowski1992role}.  The value 0.8 in this formula
was arrived at by fitting to Monte Carlo simulation.  Although
Eq.~\ref{eq:sokolowski} has the advantages of only involving density
averages at the points at which the pair distribution function is
desired, it cannot be written as a single-site convolution, since the
convolution kernel depends on both points.

@@

We construct our function's form in terms of an expansion of powers of
$h_{\sigma} = g_{\sigma}-1$, ensuring that it approaches $g_{\sigma}$
in the contact distance limit, leaving our function with tunable
constant perameters.  We run a simple monte-carlo simulation to
calculate values of the homogeneous radial distribution function at
varying distance perameter $r$.  We then tune our function's constant
parameters by matching our function's results against the monte-carlo
data, using a least squares fit method.  During this process it is
simplest to use an analytic expression for $g_{\sigma}$ in our
function and so we use the well known and very successful Carnahan
Starling equation to calculate $g_{\sigma}$.  This is appropriate
since our published $g_{\sigma}$ approaches the Carnahan Starling
version in the homogeneous limit and it's this limit, along with the
contact limit, that we would like our constructed radial distribution
function to get right.  After this process we have an appropriate
radial distribution function that is dependent only on $r$ and
$g_{\sigma}$.

The function has the form:
\begin{multline}
  g(|\rr|) = 1 + h_{\sigma}e^{-x_0(r-\sigma)} \\+ x_1h_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
  -x_4h_{\sigma}^2\sin(x_5(r-\sigma))e^{-x_6(r-\sigma)}
\end{multline}
The finalized form of our pair correlation function is:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
\end{align}

@@

In contrast to the previous works, which are based on the homogeneous
radial distribution function $g(r,\eta)$ expressed as a function of
distance and packing fraction, we have constructed our separable
approximation for the radial distribution function as $g(r,
g_\sigma)$, which is a function of the radius $r$ and its value at
contact $g_\sigma$.  One can convert between these two representation
using the Carnahan-Starling formula for $g_\sigma(\eta)$
\begin{equation}
  g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}.
\end{equation}
The form $g(r, g_\sigma)$ is more convenient for treating the
inhomogeneous case, because we can directly compute the averaged value
of $g_\sigma$ for an inhomogeneous system~\cite{schulte2012using}.

We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
\begin{equation}
  g(\sigma, g_\sigma) = g_\sigma.
\end{equation}
A second constraint applies to the integral of the total correlation
function
\begin{align}
  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
\end{align}
where $h(r) = g(r) - 1$ is the total correlation function.
A final constraint is that the slope of the radial distribution
function at contact is given to a very good approximation by
\begin{align}
  g'(\sigma) \approx - h_\sigma g_\sigma
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function.
Under these three constraints, we applied a least-squares fit to
simulation data for the radial distribution function at packing
fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
distribution function is given by
\newcommand\kappaa{\kappa_0}
\newcommand\kappab{\kappa_1}
\newcommand\kappac{\kappa_2}
\begin{multline}
  h(r,g_\sigma) = 1 + h_\sigma e^{-\kappaa z} \\
  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
  + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
      \eta^4}\frac{1}{n}
    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
\end{multline}
\begin{equation}
  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
  - \frac{4\pi}{3} \sigma^3
\end{equation}
\begin{equation}
  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
\end{equation}
\begin{equation}
  I_2 = \frac{8\pi (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
\end{equation}
\begin{equation}
  z = r - \sigma
\end{equation}
\davidsays{Can someone check the above equations for correctness
  against the python code, in plot-ghs2.py.  We'll want to
  ensure the C++ code agrees with this as well.}
This function includes three fitted parameters, $\kappaa$, $\kappab$,
and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
figure, it is apparent that the integral of $h(r)$ from
Eq.~\ref{eq:total-constraint} results in cancelling errors at large
distances.  The maximum error in $g(r)$ at packing fractions up to
0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr$.

@@

Figure~\ref{fig:pair-distribution-1}b
and~\ref{fig:pair-distribution-3}b show the pair distribution function
as plotted along a path illustrated in
Fig.~\ref{fig:pair-distribution-1}a.  These plots compare this work
with Monte Carlo results, as well as the approximations of Sokolowski
and Fischer~\cite{sokolowski1992role} and that of Fischer and
Methfessel~\cite{fischer1980born} at packing fractions of 0.1 and 0.3.
The approach of Fischer and Methfessel is only defined when the two
spheres are in contact, and is therefore only plotted on that segment
of the path.  As an input to both previous approximations we use the
hard sphere radial distribution function found with Monte Carlo
simulation, interpolated as necessary.  We find that both previous
approximations to the pair distribution function predict the
qualtitative angular dependence of the pair distribution function at
contact better than this work.  However, in each case the pair
distribution function has a systematic error at contact---either too
high or too low.  Thus while our approximation is smoother than either
of the existing approaches, its errors will have a tendancy to cancel
when used in a perturbation expansion.  At higher densities, the
approximation of Fischer and Methfessel requires evaluating the radial
distribution function at densities significantly higher than the
freezing density, which poses numerical difficulties when using the
radial distribution function from simulation.  When the two points are
more than a radius away from contact, we find that any of these
approaches gives a reasonable prediction, with our approach
underestimating the oscillations in $g^{(2)}$, as expected based on our
fit in Fig.~\ref{fig:radial-distribution}.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
  \caption{The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
    shows the triplet distribution function, with white corresponding
    to the contact value of the \emph{pair} distribution function.
    The plot on the right shows the value of $g^{(3)}$ along the path
    shown as a dotted line on the plot to the
    left.}\label{fig:triplet-contact-distribution-3}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
  \caption{The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.4, plotted when
    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
    shows the triplet distribution function, with white corresponding
    to the contact value of the \emph{pair} distribution function.
    The plot on the right shows the value of $g^{(3)}$ along the path
    shown as a dotted line on the plot to the
    left.}\label{fig:triplet-contact-distribution-4}
\end{figure}

@@

Figures~\ref{fig:triplet-contact-distribution-3}-\ref{fig:triplet-inbetween-distribution-4}
compare the triplet distribution function computed using our
approximation to $g^{(2)}$ with results from Monte Carlo simulation.
In each figure, $\rr_1$ and $\rr_2$ are held fixed, while the third
position $\rr_3$ is varied.  The test-particle position in each case
is $\rr_1$, which is on the left-hand side of the figure.  In the
color plot on the left, the first and second spheres are displayed in
grey.  On the bottom half of each left-hand plot is the triplet
correlation function as computed using the approximation presented in
this work.  The top half displays the triplet correlation function
from Monte Carlo.  On the right, we display

@@

Figure~\ref{fig:dadz-delta} shows the results for the sticky
hard-sphere fluid.  Our pair distribution function is constructed to
produce this result exactly, provided the averaged pair distribution
function at contact from Ref.~\citenum{schulte2012using} is exact.  As
expected, we see excellent agreement with the Monte Carlo simulation
results, while the approximations of Fischer and Sokolowski each show
deviations near the interface.  Figure~\ref{fig:dadz-square-well}
shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
This potential stresses our approximation to the radial distribution
function, since the edge of the attraction occurs at the dip in the
radial distribution function at this density, with poor results for
the bulk fluid.  We see that over these length scales, Sokolowski's
approximation is very accurate.  Finally, in the case of the power-law
attraction both approximations work very well.
